### 2020-08-28 B547A1/text/attack_tree.tex
moved non-repo papers from overleaf into this repo

\section{Attack Tree}
\newlist{myEnumerate}{enumerate}{9}
\setlist[myEnumerate,1]{label*=\arabic*.}
\setlist[myEnumerate,2]{label*=\arabic*.}
\setlist[myEnumerate,3]{label*=\arabic*.}
\setlist[myEnumerate,4]{label*=\arabic*.}
\setlist[myEnumerate,5]{label*=\arabic*.}
\setlist[myEnumerate,6]{label*=\arabic*.}
\setlist[myEnumerate,7]{label*=\arabic*.}
\setlist[myEnumerate,8]{label*=\arabic*.}
\setlist[myEnumerate,9]{label*=\arabic*.}

\begin{myEnumerate}

\item Compromise the software/firmware update
  \begin{myEnumerate}[label*=\arabic*.]
  \item Gather knowledge (and)
    \begin{myEnumerate}[label*=\arabic*.]
    \item Find the IP address or URL of the webserver (or)
      \begin{myEnumerate}[label*=\arabic*.]
      \item Google the website of the target USB
      \item Collect the static contents of the web site (e.g. CSS, javascript,
        images, \dots).
      \end{myEnumerate}
    \item Get the same model as the target USB
      \begin{myEnumerate}[label*=\arabic*.]
      \item Try to see if it is easy to feed in any malware/keylogger
      \end{myEnumerate}
    \item Find the place (IP range) where he/she uses the target USB
      \begin{myEnumerate}[label*=\arabic*.]
      \item Use port scanner in the same subnet
      \item Tap his/her network cable
      \item Find the public key of the target USB
      \end{myEnumerate}
    \end{myEnumerate}
  \item Build a fake web site which looks like the real one (and)
    \begin{myEnumerate}[label*=\arabic*.]
    \item Prepare the MITM attack
      \begin{myEnumerate}[label*=\arabic*.]
      \item Do the reverse engineering to build a malicious app to feed into the
        target USB
      \end{myEnumerate}
    \end{myEnumerate}
  \item Gain an access (and)
    \begin{myEnumerate}[label*=\arabic*.]
    \item Use the DNS Spoofing trick to redirect the victims access of app
      update to the fake website
      \begin{myEnumerate}[label*=\arabic*.]
      \item Wait until the target USB tries to access the website to get it updated (or)
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \item Mail a fake notification of the new release of the app with the URL of
        the fake website
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \end{myEnumerate}
    \item Use the IP Spoofing trick to intercept the victims access of app
      update to the fake website
      \begin{myEnumerate}[label*=\arabic*.]
      \item Wait until the target USB tries to access the website to get it updated (or)
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \item Mail a fake notification of the new release of the app with the URL of
        the fake website
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \end{myEnumerate}
    \end{myEnumerate}
  \end{myEnumerate}
\end{myEnumerate}