### 2020-08-28 B547A1/text/introduction.tex

\chapter{Introduction}
\label{ch:introduction}

This report satisfies the requirements of assignment
1~\cite{2017myersa1handout}, which requires us to perform a
\textit{comprehensive} threat model of a password manager.  We describe the
password manager, its functionality and provide state, swimlane and dataflow diagrams
which we use in our threat model. We also perform risk analysis, describe
mitigations in the threat model and provide attack trees for several of our
highest rated threats.  \sidenote[][-0.5in]{Our password manager is hosted on a
specialized USB device which emphasizes hardware functionality as a means of
providing acceptable protection to stored passwords.}

\par We threat modelled using the STRIDE approach, focusing on data
flows and security-relevant elements.
%\sidenote[][]{STRIDE: \emph{Spoofing},\emph{Tampering},\emph{Information Disclosure},\emph{Denial of Service}, and \emph{Elevation of Privilege}}
Similarly, we manage our threats and risks using
Shostack~\cite{shostackbook} and modified OWASP
methodologies~\cite{owasprisk}.

%We are allowed discretion in the choice of architectures but the architecture
%should provide for the remembering of many passwords for websites,
%products, appliances and other uses which require passwords.

%\par The report is limited to no more than 7 pages of diagrams and 40
%pages total length.  The final product is a threat model of the system
%and any other documentation that would be necessary to understand how
%the proposed system would run.  If any sections of the threat model
%are empty, they should still be included in the report and the report
%should indicate that they were intentionally left empty.