### 2020-08-28 B547A1/text/threat_list.tex

\cleardoublepage
\chapter{Appendix 2: Threat List}
\label{ch:threatlist}

The final system threat table is shown in figure~\ref{tab:threatlist}.  Each
entry in the threat table has an associated ID, who found the threat, the DFD
diagram element which applies to the threat, the type of threat as given by the
STRIDE model, and a mitigation.  Each entry is also assign a bug ID and finally
a determination of whether the entry is accepted into the larger threat risk
table.  \marginnote{In practice, the use of design rules can be very effective
because they form tangible requirements which are push down to the lowest level
and then reviewed as the software product matures at each design review.  This
keeps key security practices up front and visible throughout the design process.
Finally,incorporating design rules into an evaluation plan ensures the integrety
and security of the final product.}
\par If one threat is similar to another, it may not be entered into the threat
table but rather a threat risk that best decstibes the collective threat is
entered into the table.  This does not mean that the threat is ignored as it
still has a bug id associated with it which must be dispositioned.  In some
cases the threat is not carried to the risk table but instead may be referred to
a design rule which is then checked off during software component design reviews
and testing.

\begin{table*}
\centering
  \includegraphics[width=\linewidth]{threat_list_final}
  \caption{Final Table of Threats}
  \label{tab:threatlist}
\end{table*}