### 2020-08-31 CyberKGX/body.tex
cleaned old paper info from CyberKGX files

\section{Introduction}
TBD

\paragraph{Our contributions.} TBD

\section{Methodology}

\subsection{Computing Environment}

TBD

\subsection{Data}

TBD

\subsection{Models}

TBD

\subsection{Task}
TBD

\section{Results}

TBD

\section{Conclusions}

TBD

\section{Future Work}

TBD

\begin{acks}

This research was supported in part by Lilly Endowment, Inc., through its support for the Indiana University Pervasive Technology Institute. \cite{PTI} % for Carbonate use

TBD - remember compute resources and other boilerplate acks

\end{acks}

\newpage
\appendix
%\newpage
%Appendix A
\section{Appendix A Title}

TBD

%Appendix B
\section{Appendix B Title}

TBD